
\chapter{The liquid crystalline phases}
\label{chap:one}

\introduction

Up to the end of the $19^{\mathrm{th}}$ century, the three known states of matter were gas,
liquid and solid. According to dictionary definitions~\cite{physicsDictionary, structMatter} 
a substance is gaseous if it expands to fill its container homogeneously, regardless of its volume, 
displays a high compressibility and shows no positional correlations on length scales greater than
the molecular size. At the other extreme, a solid retains its shape, can support shear and its
atoms are restricted in space, oscillating about a fixed position. Two type of solids
can be distinguished; in crystalline solids, the atoms are regularly spaced on a 3 dimensional
lattice, whereas in amorphous or glassy solids, the atoms are disordered on a large scale
but ordered in the range of a few molecular lengths.\\
Liquid is a state of matter which exists between the two preceeding phases. A liquid substance 
only fills part of its container and its localisation is largely controlled by the gravitational 
force. Actually a liquid is only one component of a two phase system as every liquid is always 
searching to be in equilibrium with its own saturated vapor. A liquid has no rigidity, gives 
no resistance to shear under static conditions and has a small volume compressibility. On a 
microscopic scale, the positions of the constituent molecules are randomly distributed and no 
long range order can be found. Molecules in a liquid are subject to Brownian diffusion.\\

However, this picture changed somewhat in the late eighteen eighties. Slightly before then a number 
of scientist noticed some uncommon crystallization behaviour in certain substances which were
found to  transform from isotropic
liquid to a non-crystalline form before undergoing full crystallization. At the time, this
was attributed to the presence of impurities in the samples. The actual discovery of liquid
crystals is attributed to Friedrich Reinitzer~\cite{Reinitzer88} who, in 1888, was studying 
a compound related to cholesterol (cholesteryl benzoate). Reinitzer observed what he referred 
to as `two melting points' and identified the new phase; this was later termed liquid crystal 
by his friend and colleague Otto Lehmann~\cite{Lehmann89} who performed the first polarised 
optical microscopic measurements on liquid crystals.\\
Although not much studied at first because of the lack of the direct applications, 
interest in liquid crystals increased dramatically during the $20^{\mathrm{th}}$ century, mostly 
because of Liquid Crystal Display (LCD) applications; a wide understanding of these phases and 
the molecules that form them has now been built up.



%===================================================================================================
%===================================================================================================
\input{src/11-chap1/generalLC}
\input{src/11-chap1/flexoLC}
\input{src/11-chap1/LCtheories}
\input{src/11-chap1/expLC}
\input{src/11-chap1/applicationLC}
%===================================================================================================
%===================================================================================================




