

\section{Conclusions}

%==============================================================
%		SYMMETRIC CONFINED SYSTEMS
%==============================================================
The study of confined hard particle liquid crystalline systems has been addressed 
in Chapters~\ref{chap:four} and~\ref{chap:five}. The Surface induced structural 
changes in symmetric anchored systems have
been studied using three surface interaction models, namely the hard needle wall (HNW),
rod-sphere (RSP) and rod-surface (RSUP) potentials.  For all three of these models, the surface
interaction was not mediated directly by the HGO particles, but by other objects embedded
within them. This approach allowed the study to be extended to include consideration of 
the effect of varying particle
absorption into the surface. Through this, it has been shown that the HNW and RSUP potentials 
induce planar
and homeotropic anchoring for, respectively, long and short elongations of the inner object
($k_S$) whereas with the RSP potential and long $k_S$, the planar arrangement was replaced by 
tilted anchoring. These results have proved to be consistent with a theoretical
treatment based on the geometrical characteristics of the surface interaction models.
The implications of these results are twofold. The behaviour found for the HNW and RSUP models
showed that the anchoring transition between
homeotropic and planar arrangements can be controlled by the molecular volume made available to
be absorbed into the substrates. Secondly, the tilted phase obtained with the RSP model lead to
a revision of the
explanation of tilted surface arrangements which had previously always been attributed to
attractive forces. The observation of such a structure in this study showed that it can also be
obtained with a purely steric potential.\\
Anchoring and order phase diagrams have been computed for all models. For the HNW
and RSUP potentials, using series of simulations at constant densities and either increasing or
decreasing $k_S$, regions of bistability between the planar and homeotropic surface arrangement
have been found. Those regions proved to be stronger and wider for the RSUP model. 
This shows that even in very simple model systems, bistable surface behaviour can be introduced
by tuning the competition between two locally stables states. This suggests that similar tuning
of real liquid crystal substrate systems may offer a viable route to relatively simple bistable
surfaces for low energy device applications.\\

Hybrid anchored systems with a top homeotropic surface have been studied in
Chapter~\ref{chap:five} using the HNW as a surface potential in an attempt to create a hybrid
anchored slab with a top surface homeotropic anchoring and bistable anchoring at bottom
substrate. It has been shown that the bistability behaviour of the model at the bottom surface
is lost if the top homeotropic anchoring is made too strong. However, by reducing the latter
(\ie reducing $k_S$), the bistability at the bottom surface can be recovered. This has been
explained in terms of the elastic forces transmitted from the top surface particles onto those at the
bottom surface. Reducing the top anchoring strength reduces these forces and, thus, 
allows bistability to be recovered. The point has also been addressed in Chapter~\ref{chap:seven}.
Also, the behaviour of the homeotropic to planar structural transition has been shown to be
dependent upon the slab height. For short slabs (\ie small systems), there is a discontinuous
structural transition between the two arrangements, whereas this becomes continuous for thicker
slabs (\ie bigger systems). These findings are consistent with both
experimental and theoretical studies of the effect of thickness on the bent director
structure of a hybrid anchored system. Also, this result has shown that the first prerequisite
for the modeling of a bistable LCD cell can be met: it is possible to simulate a hybrid
anchored slab with a homeotropically anchored top surface and bistable anchoring at the bottom
surface using an appropriate parameterisation for the surface model. In addition, provided large
enough systems are used, a continuous structural transition from homeotropic to planar alignment
can be obtained if the slab is in the HAN configuration.\\


%==============================================================
%		PEAR SHAPED PARTICLES
%==============================================================
Chapter~\ref{chap:six} has addressed the study of tapered pear-shaped particles. The first model to
be used for this was the truncated Stone expansion model; this, however, did not meet the 
requirement of
displaying a stable nematic phase and showed only two phases: an isotropic phase at low density
and glassy domain-ordered phase at high density. This behaviour proved to be very surprising as
an equivalent soft model used by the Bologna group showed isotropic, nematic and \smA phases. The
lack of ordered fluid phases in the steric model was attributed to concave features in
the particles' contact surfaces which had the effect of preventing the particles from sliding along
one another.
As an alternative, the PHGO model, a generalisation of the HGO model to non-centrosymmetric particles
has been developed. This does not show the concave features mentioned above. For the two shortest
elongations used, $k=3$ and $4$, the PHGO model only showed isotropic and fluid domain-ordered
phases at, respectively, low and high densities. However, upon increasing the elongation to $k=5$,
the model phase behaviour became much richer, with transitions from isotropic to nematic and
then to a bilayered smectic $\mrm{A_2}$ phase. The study of this latter
phase proved to be very interesting, showing anisotropic compressibility behaviour. Thus the
PHGO model, with its nematic phase, met the second requirement for the bistable LCD cell
model.\\


%==============================================================
%		FLEXOELECTRIC SWITCHING
%==============================================================
In Chapter~\ref{chap:seven}, confined systems of PHGO particles interacting with the substrates
through a variant of the RSUP potential have been studied. This surface potential has been shown to
exhibit both planar and homeotropic arrangements with a narrow region of bistability around
$k_S/k=0.7$. In the case of hybrid anchored systems of $N=2000$ particles with a top homeotropic
surface, bistability at the bottom surface could be recovered using $k_{St}/k=0.6$. Due
to the large system size used, a continuous structural transition was found in the case of the bent
director state.\\ 
Using this, a molecular model for the LCD cell was simulated and both hard and easy 
switching between the HAN and vertical states were attempted using both dielectric and dipolar
contributions in the particle-field interactions. Both switching directions could be achieved 
using a narrow window of electric field strength $E$ and dipole moment $\mu$. A successful
parameterisation was $E/\delta\epsilon = \pm 0.2$ and $\mu/\delta\epsilon = 2.5$. However the
question as to  whether the mechanism underlying this switching behaviour is direct surface
effect, indirect flexoelectric behaviour or a combination of the two has not been resolved; 
this forms, in part, the basis of future work.







