
%===============================================================================================
%			NOTES
%===============================================================================================
%
%===============================================================================================


\chapter{Computer simulations of Hard Gaussian Overlaps}
\label{chap:three}


\introduction

In the previous Chapter it was shown that the Gay-Berne potential is one of the most versatile 
molecular models for liquid crystal simulations. Depending on the chosen shape and energy 
parameterisation, it can be used to model many liquid crystalline phases. 
However, when the nematic phase only is of interest, the use of
hard particle models is a preferable option, being computationally easier and faster to
implement than the soft models. Also hard particles have proven to be a very good test-bed
for perturbation theories.\\
In this Chapter the computer simulation of such a model, the hard Gaussian overlap (HGO) 
model is discussed.  First a literature review of the computer simulations and theoretical 
work performed on  this model is provided. The techniques used for the computation of the most
relevant observables in computer simulations are then presented. Finally, some preliminary
results from bulk simulations of prolate and oblate HGO particles 
are presented.  A comparison with existing results for calamitic molecules is also included.

%=================================================================
\input{src/13-chap3/HGO}
\input{src/13-chap3/observables}
\input{src/13-chap3/HGOresults}
%=================================================================
