

\chapter{Surface influence on liquid crystalline systems}
\label{chap:four}

%===============================================================================================
%===============================================================================================
\introduction

When a Liquid Crystal is placed in contact with another phase, a surface is created which breaks
the symmetry of the system. There are two main effects to this symmetry breaking~\cite{Jerome91}.
The first is to introduce molecular layering close to the interface while the second  is to 
modify the surface orientational ordering in the interfacial region.\\
The purpose of this Chapter is twofold. First, the effects of confinement are studied on systems of
ellipsoidal particles represented by the hard Gaussian overlap model and using a simple 
particle-surface potential, the so called Hard Needle Wall potential. The effects of varying
density and particle-substrate interactions are studied.\\
The second aim of this Chapter is to study the anchoring transition from planar to homeotropic
alignment. As very little simulation work has been performed on anchoring transitions, it is
hoped that this study can shed some light on their origins. Finally, possible
bistability regions in the anchoring behaviour of the models are examined.


%===============================================================================================
%===============================================================================================

\input{CHAP4/HNWmodel.tex}
\input{CHAP4/symmetricResults}
\input{CHAP4/surfInfPhaseTrans}



%===============================================================================================
%===============================================================================================
\conclusion

In this Chapter, the study of surface induced structural changes on a confined system of hard
Gaussian overlap particles has been addressed. The choice of the hard needle wall potential for
surface interaction allowed the observation of two stable surface arrangements, namely planar
and homeotropic according to $k_S$, the length of the needle embedded in the particles. The
mechanism responsible for the change in surface arrangement as a function of $k_S$ is the
varying amount of molecular volume that can be absorbed into the surface.\\
A systematic study of the behaviour of the system as a function of number density and needle 
length has been performed. From
this, an anchoring transition between the two arrangements has been identified and located
through the computation of anchoring phase diagrams. Also, differences between the diagrams
obtained from series of simulations performed with, respectively increasing and decreasing 
needle lengths have been used to identify bistability regions, where both homeotropic 
and planar arrangement remain stable on timescale of a simulation run length.\\ 
It has been shown that generally, the effect of confinement is to shift the isotropic-nematic 
transition to lower number densities close to the surfaces and to higher number densities 
in the bulk region.
In addition, it has been shown that the amount by which this transition is shifted varies with
the nature of anchoring conditions adopted.\\


%===============================================================================================
%===============================================================================================
