
\chapter{Computer simulations of liquid crystals}
\label{chap:two}

\introduction

Computer simulations act as a link between theory and experiment. Experimental analysis is
limited in that, usually,  only bulk averaged properties are measurable meanwhile, the complicated
nature of mesogens and their interaction potentials prevents the possibility of a full theoretical 
treatment capable of  predicting the phase behaviour of real mesogens. These
limitations can be resolved using computer simulations, where, by using a model interaction
in an appropriately designed `computer experiment', full molecular insight into the system can be
gained. Using the results of statistical mechanics, macroscopic properties can be computed 
which, in turn, can be used to test the theories from which the model originated.
The validity of the model can, thus, be checked against experimental results. Computer simulations can
also act as predictive tools in the design of novel compounds or for the study of systems under
conditions that can only be attained in the laboratory with great difficulty and cost.\\
The first part of this Chapter contains a description of the two main methods used in the simulations 
of liquid crystals at a
molecular scale, namely the Monte Carlo method and Molecular dynamics. Following this, previous
work on the computer simulation of liquid crystals is reviewed. Here attention is concentrated on
generic hard particles and Gay-Berne models, and their extension to treat flexoelectric
behaviour.
The Chapter then concludes with a review of progress on the simulation of confined liquid 
crystalline systems and their anchoring properties.


%==============================================================================
%==============================================================================
\input{src/12-chap2/modelingTechs}
\input{src/12-chap2/models}
\input{src/12-chap2/flexoModel}
\input{src/12-chap2/confinedLC}
\input{src/12-chap2/anchoringTransitions}
%==============================================================================
%==============================================================================
