

\chapter{More on confined geometries}
\label{chap:five}
%===============================================================================
%===============================================================================

\introduction

Simulations of confined systems of hard Gaussian overlap particles interacting with the
substrates through the hard needle wall potential have shown that, despite the simplicity 
of the model, a range of surface induced behaviour can be observed. Using this simple setup 
a thorough and systematic study of the surface induced structural changes has been performed 
and an anchoring transition has been identified. In this Chapter, the focus is 
brought to bear on the study of alternative confined system configurations.\\
%
First, more realistic surface potentials are studied and their anchoring phase diagrams computed 
so as to investigate their suitability for the modeling of anchoring transitions. Two such 
potentials are of interest, namely the rod-sphere potential and the rod-surface potential. For
each of these, the the surface induced arrangements are studied and the possibility of 
bistable regions is explored.\\
%
The second part of this Chapter contains a study of hybrid anchored systems performed
using the hard needle wall potential; following this, the possibility of simulating 
easy switching between the hybrid aligned nematic and vertical states of the cell is investigated.


%===============================================================================
%===============================================================================
\input{src/15-chap5/realisticPotentials_RSP}
\input{src/15-chap5/realisticPotentials_RSUP}
\input{src/15-chap5/hybridSystems}

%===============================================================================

\conclusion

In this Chapter, two issues have been addressed. In the first part of the Chapter, two
more realistic surface potentials have been studied, namely the rod-sphere and the rod-surface
potentials. The aim of this work was to find a potential which has a more realistic basis
than the hard needle wall potential but which also displays planar and homeotropic surface
arrangements. A region of bistability between the two arrangement was also required for future
applications relating the modeling of display cells.\\
The rod-sphere potential was found to be unsuitable as the planar arrangement was replaced by a
tilted structure. However, the results obtained using this model proved to be interesting
since they showed that a tilted phase can be obtained from purely steric interactions. The
rod-surface potential, meanwhile, proved a better candidate for the aim stated above, as it
not only recovered the surface behaviour of the HNW potential but actually displayed stronger
and wider bistability regions.\\
In the second part of this Chapter, hybrid anchored systems of HGO particles confined between a
homeotropic top surface and a bottom surface with competing anchoring have been studied using
the HNW potential for the surface interactions. These simulations showed that the
bistability behaviour of the model can be lost if high anchoring strength is used at the top
surface or if the slab is too narrow. Using moderate homeotropic anchoring at the top surfaces
and systems sizes of $N=2000$ particles, however, bistability was established and a cell
was successfully switched between the HAN and V states if only the easy switching was
considered. Achievement of the reverse switching requires the use of flexoelectric particles and
a surface potential allowing bistability between planar and homeotropic arrangements for such
particles. Those two problems are addressed, respectively, in Chapters~\ref{chap:six}
and~\ref{chap:seven}.






