

\chapter{Bulk simulations of pear shaped particles}
\label{chap:six}

%===================================================================================================
%===================================================================================================
\introduction

The last two Chapters have dealt with surface induced effects in systems of hard Gaussian 
ellipsoids subject to a range of different surface potentials. For some of these, bistable regions 
corresponding to different surface arrangements have been found. Here, a new line of work is
considered, the focus of which the behaviour of non-centrosymmetric, pear-shaped particles.
In the next Chapter, these two threads will be brought together in a study of these pear-shaped
particles in confined geometries.\\
It is recalled that the ultimate aim of this thesis is to model surface induced switching between
two arrangements of an hybrid anchored cell following the theoretical
treatment of Davidson and Mottram~\cite{DavidsonMottram02}. The system properties required
to achieve this 
are a bistable surface and a stable nematic phase of flexoelectric particles. The former problem
has already been addressed; this Chapter deals with the latter. For this, the development of a
model for pear-shaped particles is studied as particles of that shape are thought to display 
flexoelectric behaviour~\cite{Meyer69}.\\
In this Chapter, two models are studied, the first one hereafter referred to as the HP model is a
hard version of the potential  used in~\cite{BerardiRicci01}. Despite having a rich phase
behaviour, 
we show that this model does not have a stable nematic phase. As a result a second model,
which we term the Parametric Hard Gaussian Overlap (PHGO) model, has been developed which does
yield the required phase behaviour. The work presented 
in this Chapter has been submitted for publication to Physical Review E~\cite{BarmesRicci03}.

%===================================================================================================
%===================================================================================================
\input{CHAP6/HPmodel}
\input{CHAP6/HPresults}
\input{CHAP6/PHGOmodel}
\input{CHAP6/PHGOresults}


%===================================================================================================
%===================================================================================================
\conclusion


In this Chapter the simulation of non-centrosymmetric, pear-shaped particles has been addressed
using two models. The first model used a steric version of the model developed by 
Berardi~\etal\cite{BerardiRicci01}.
The contact distance of this is obtained fitting a numerical contact distance
to a truncated Stone expansion. However, this introduced non-convex contact surfaces and, 
as a result, the model with
elongations $k=3$ and $5$ did not display stable nematic phases as they were systematically 
preceeded by glassy phases. A totally new approach was then taken in the form of the  so called
parametric hard Gaussian overlap model. For the shorter elongations $k=3$ and $4$, the model
did not show any nematic phases; rather at high densities, the molecules order into
interdigitated domains. 
Upon increasing the elongation to $k=5$, the model was found to display both
nematic and interdigitated smectic $\mrm{A_2}$ phases. The latter showed some interesting
anisotropic compressibility behaviour where compression had the effect of decreasing the
intra-layer particle separation while increasing the distances between bilayers. As a result, 
the PHGO model with an elongation $k=5$ can be used for the modeling of surface induced 
switching in hybrid anchored systems of flexoelectric particles; the implementation of this 
is discussed in the next Chapter.






