

%\vspace*{5mm}
\hrule
\begin{center}
	\textbf{\large Acknowledgments}
\end{center}
%\vspace*{5mm}
\hrule
\vspace*{15mm}


Over the past three years, I have met and spent time with many people who directly or indirectly
made their contribution to this Thesis and who I am indebted to.\\
%
%===================================================
%	WORK 
%===================================================
First of all, I would like to thank
my supervisors in Sheffield, Dr. Douglas J. Cleaver and Pr. Christopher M. Care for giving me 
the opportunity of taking this PhD and for their guidance and help throughout this project. 
A big \emph{gracie} also to my third supervisor, Pr. Claudio Zannoni for offering me the
opportunity of conducting part of my research in his group in Bologna. Also, my work during the
past three years would not have been that enjoyable without the presence of my many colleagues
in Sheffield, Mike, David, Dmytro, Fatima, Vicktor, Richard, Rachael, Kev, Lee, Sergei, Lloyd,
Gavin, Balaczs, Tim and those in Bologna, Matteo, Davide, Luca, Silvia, Bebo, Alberto,
Corrado, Tim and Alberto. Thanks also to Kevin Blake for his expertise as a system administrator and
all his guidance which helped me so much in my battle with Linux and Unix systems. I must also
mention the valuable help from Dmytro in deriving the results shown in Appendix~\ref{chap:A}.\\
%
Also, I can not forget all those people who helped me in having a life out of the lab. Especially
a big thank you to Steve and Serena who made my induction to this country so easy and enjoyable.
I have spent a brilliant time living at their place. Thanks also to the `corps' at the
Hallamshire aikido club, Lou, Will, Ben, Mark, Adam, Steve, Gavin, Simon and Chris who make the
standard of aikido in Sheffield something to be reckoned with. 
Thanks also to Dmytro, Ben and Mike for being such fine housemates.\\
%
Last but by far not least, I wish to acknowledge my family, especially my parents, my
sister and Audrey for their love and support.

