\newpage
\hrule
\begin{center}
	\textbf{\large Advanced studies}
\end{center}
%\vspace*{5mm}
\hrule
\vspace*{10mm}

\begin{itemize}

\item September 2000~: Complex fluids 2000, Sheffield Hallam University.

\item December 2000~: 
	British Liquid Crystal Society (BLCS) Winter Workshop. University of Hull.
	
\item July 2001~: 
	Collaborative Computational Project 5 (CCP5) summer school, methods in molecular simulations. 
	UMIST.
	Poster presentation \emph{``A computer simulation study of modified surface interactions in
	liquid crystalline systems''}
	
\item July 2001~: 
	presentation \emph{``A computer simulation study of modified surface interactions in
	liquid crystalline systems''}.
	Materials Research Institute, Sheffield Hallam University.

\item September 2001~: 	
	Poster presentation \emph{``A computer simulation study of modified surface interactions in
	liquid crystalline systems''}.
	CCP5 annual meeting, ``Liquids and Liquid Interfaces'', Warwick university.

\item  May 2002~: 
	presentation \emph{``Surface induced structural changes in confined liquid
	crystalline systems: a computer simulation study''}. Univerist\'a di Bologna.

\item September 2003~: 
	presentation \emph{``Towards modeling surface induced switching in bistable devices''}.
	CCP5 annual meeting, ``Advances in simulations of molecules and materials'', Durham university.

\item February 2003~: 
	presentation \emph{``Towards modeling surface induced switching in bistable devices''}. 
	Materials Research Institute, Sheffield Hallam University.

\item July 2003~:
	presentation \emph{``Flexoelectricity induced switching in bistable LCD cells''}. 
	University of Strathclyde.


\end{itemize}


